Synpotic Soil Characterization

### 2024-11-13 3-reports/manuscript_figures.Rmd
recode region and transect

corr_all = fit_correlations_function(dat = data_wide_PCA)
#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "WLE"), TITLE = "WLE" )
#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "CB"), TITLE = "CB")

@@

```{r pca_input}
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
```

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

@@

pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "CB")) %>% mutate(region = "CB")

@@

```{r scaled_gg1, fig.height = 8, fig.width = 7}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r v2-pca_input}
pca_overall = fit_pca_function(data_wide_PCA2) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "CB"))
```

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wc"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wc"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-5, 3)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1,
                                    labels = c("A: WLE", "B: CB"))
gg_pca_regions

@@

pc_overall = get_pc_values(data_wide_PCA2) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "CB")) %>% mutate(region = "CB")

@@

```{r v2-scaled_gg1, fig.height = 8, fig.width = 7, eval=FALSE}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r v2-scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance, SOM, DIC, Al)) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r analytesx-2, fig.height = 4, fig.width = 8, eval=FALSE}
data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = region), alpha = 0.6)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  facet_wrap(~region)+
#  scale_fill_manual(values = c("darkblue", "red"))+
  scale_fill_manual(values = pnw_palette("Bay", 2))+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")
```

@@

  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
    filter(!is.na(value))

@@

# plot Specific Conductance
gg_spc =
  data_combined_subset %>% filter(grepl("spCond", name)) %>% filter(region == "CB") %>% 
  ggplot(aes(x = transect, y = value)) +
    # plot points
    geom_point(aes(group = site, color = site),
               size = 2.5, stroke = 1, position = position_dodge(width = 0.4))+
    scale_alpha_manual(values = c(1, 0.3))+
    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                       values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
  #  facet_wrap(~region, labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
    theme_kp()+
    labs(x = "",
         y = "Specific Conductance, mS/cm",
         shape = "Site",
         color = "Site")+
    theme(legend.position = c(0.2, 0.8), legend.box = "vertical")+
    NULL

# plot Chloride
gg_cl = 
  data_combined_subset %>% filter(grepl("Chloride", name)) %>% filter(region == "CB") %>% 
  ggplot(aes(x = transect, y = value)) +
    # plot points
    geom_point(aes(group = site, color = site),
               size = 2.5, stroke = 1, position = position_dodge(width = 0.4))+
    scale_alpha_manual(values = c(1, 0.3))+
    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                       values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
  #  facet_wrap(~region, labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
    theme_kp()+
    labs(x = "",
         y = "Chloride, meq/100g",
         shape = "Site",
         color = "Site")+
    theme(legend.position = c(0.2, 0.8), legend.box = "vertical")+
    NULL

>>>>

corr_all = fit_correlations_function(dat = data_wide_PCA)
#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "Erie"), TITLE = "Erie" )
#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "Chesapeake"), TITLE = "Chesapeake")

@@

```{r pca_input}
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "Erie"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "Chesapeake"))
```

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: Erie",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: Chesapeake",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

@@

pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "Erie")) %>% mutate(region = "Erie")
pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "Chesapeake")) %>% mutate(region = "Chesapeake")

@@

```{r scaled_gg1, fig.height = 8, fig.width = 7}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r v2-pca_input}
pca_overall = fit_pca_function(data_wide_PCA2) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "Erie"))
pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "Chesapeake"))
```

@@

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=5,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 fill = groups), 
             stroke = 1)+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                     values = pal_transect)+
  scale_fill_manual(breaks = c("upland", "transition", "wte", "wetland"), 
                    values = pal_transect)+
  scale_shape_manual(values = c(21, 24, 22))+
  xlim(-5, 3)+
  ylim(-2.5, 3.5)+
  labs(shape="", fill = "", color = "")+ 
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  guides(fill = guide_legend(override.aes = list(alpha=1, shape = 21, size = 3)))+
  NULL

gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1,
                                    labels = c("A: Erie", "B: Chesapeake"))
gg_pca_regions

@@

pc_overall = get_pc_values(data_wide_PCA2) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "Erie")) %>% mutate(region = "Erie")
pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "Chesapeake")) %>% mutate(region = "Chesapeake")

@@

```{r v2-scaled_gg1, fig.height = 8, fig.width = 7, eval=FALSE}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r v2-scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance, SOM, DIC, Al)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

@@

```{r analytesx-2, fig.height = 4, fig.width = 8, eval=FALSE}
data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = region), alpha = 0.6)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  facet_wrap(~region)+
#  scale_fill_manual(values = c("darkblue", "red"))+
  scale_fill_manual(values = pnw_palette("Bay", 2))+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")
```

@@

  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
    filter(!is.na(value))

@@

# plot Specific Conductance
gg_spc =
  data_combined_subset %>% filter(grepl("spCond", name)) %>% filter(region == "Chesapeake") %>% 
  ggplot(aes(x = transect, y = value)) +
    # plot points
    geom_point(aes(group = site, color = site),
               size = 2.5, stroke = 1, position = position_dodge(width = 0.4))+
    scale_alpha_manual(values = c(1, 0.3))+
    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                       values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
    theme_kp()+
    labs(x = "",
         y = "Specific Conductance, mS/cm",
         shape = "Site",
         color = "Site")+
    theme(legend.position = c(0.2, 0.8), legend.box = "vertical")+
    NULL

# plot Chloride
gg_cl = 
  data_combined_subset %>% filter(grepl("Chloride", name)) %>% filter(region == "Chesapeake") %>% 
  ggplot(aes(x = transect, y = value)) +
    # plot points
    geom_point(aes(group = site, color = site),
               size = 2.5, stroke = 1, position = position_dodge(width = 0.4))+
    scale_alpha_manual(values = c(1, 0.3))+
    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                       values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
    theme_kp()+
    labs(x = "",
         y = "Chloride, meq/100g",
         shape = "Site",
         color = "Site")+
    theme(legend.position = c(0.2, 0.8), legend.box = "vertical")+
    NULL
